`treats`: a modular `R` package for simulating trees and traits.

### 2023-11-22 inst/gitbook/_book/treats_manual.tex
change time.condition to age.condition
For bypassing s3 issues on the CRAN

\title{treats manual}
\author{Thomas Guillerme (\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}})}
\date{2023-11-21}

@@

\begin{verbatim}
##        --- dispRity package ---
## This is the CRAN release version (1.7.0) of the package.
## For news, vignettes and future releases,
## visit https://github.com/TGuillerme/dispRity
\end{verbatim}

@@

\begin{verbatim}
## $main
## $main$A
## $main$A$process
## function (x0 = 0, edge.length = 1, Sigma = diag(length(x0)), 
##     ...) 
## {
##     return(t(MASS::mvrnorm(n = 1, mu = x0, Sigma = Sigma * edge.length, 
##         ...)))
## }
## <bytecode: 0x5647036c6610>
## <environment: namespace:treats>
## 
## $main$A$start
## [1] 0
## 
## $main$A$trait_id
## [1] 1
## 
## 
## 
## $background
## NULL
\end{verbatim}

@@

\begin{verbatim}
## $waiting
## $waiting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(rexp(1, sum(lineage$n * (bd.params$speciation + bd.params$extinction))))
## }
## <bytecode: 0x564707510428>
## <environment: namespace:treats>
## 
## $waiting$internal
## NULL
## 
## 
## $selecting
## $selecting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(sample(lineage$n, 1))
## }
## <bytecode: 0x5647075127a8>
## <environment: namespace:treats>
## 
## $selecting$internal
## NULL
## 
## 
## $speciating
## $speciating$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(runif(1) <= (bd.params$speciation/(bd.params$speciation + 
##         bd.params$extinction)))
## }
## <bytecode: 0x56470751e620>
## <environment: namespace:treats>
## 
## $speciating$internal
## NULL
## 
## 
## $call
## $call$waiting
## $call$waiting$fun
## [1] "default"
## 
## 
## $call$selecting
## $call$selecting$fun
## [1] "default"
## 
## 
## $call$speciating
## $call$speciating$fun
## [1] "default"
\end{verbatim}

@@

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.19\columnwidth}\raggedright
function\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright
arguments\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright
input\strut
\end{minipage} & \begin{minipage}[b]{0.33\columnwidth}\raggedright
what does it do\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.19\columnwidth}\raggedright
\texttt{time.condition}, \texttt{taxa.condition}, \texttt{trait.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
the value to check the target against\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
e.g.~time = x, taxa = x or trait = x (see below)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright
condition\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
a relational operator\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
e.g.~\texttt{==}, \texttt{\textgreater{}}, \texttt{!=}, etc. for asking for example \texttt{time\ ==\ x}\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright
\ldots{}\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
any additional arguments to be passed to the event\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{itemize}
\tightlist
\item
  \texttt{time.condition} will trigger the \texttt{event} once a certain time is reached in the simulations and is the simplest/most basic condition with no arguments other than the time required. For example \texttt{time.condition(4,\ condition\ =\ \textasciigrave{}==\textasciigrave{})} will trigger the \texttt{event} once the simulations reaches 4 time units. Easy.
\item
  \texttt{taxa.condition} will trigger the \texttt{event} once a certain number of taxa are reached. This can be considered including or excluding fossil species. For example \texttt{taxa.condition(42,\ condition\ =\ \textasciigrave{}\textgreater{}=\textasciigrave{},\ living\ =\ TRUE)} will trigger the \texttt{event} if there are at least 42 living taxa.
\item
  \texttt{trait.condition} will trigger the \texttt{event} once a certain trait value is reached. This function allows to say which trait(s) to target (by default, the first one using \texttt{trait\ =\ 1}), what value of the trait to target (by default \texttt{what\ =\ max}) and whether to use and absolute trait value or not (\texttt{absolute\ =\ TRUE}). For example \texttt{trait.condition(1/3,\ condition\ =\ \textasciigrave{}\textgreater{}\textasciigrave{},\ trait\ =\ 1,\ what\ =\ sd)} will trigger the condition after the standard deviation of the first trait reaches 1/3.
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# 80\% mass extinction at time 4}
\NormalTok{mass\_extinction \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                      \DataTypeTok{target =} \StringTok{"taxa"}\NormalTok{,}
                      \DataTypeTok{condition =} \KeywordTok{time.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                      \DataTypeTok{modification =} \KeywordTok{random.extinction}\NormalTok{(}\FloatTok{0.8}\NormalTok{))}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmptzertD/file14f8c6a285c1a}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmptzertD/file14f8c301a0acb}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.55\columnwidth}\raggedright
Functionality showcase\strut
\end{minipage} & \begin{minipage}[b]{0.39\columnwidth}\raggedright
Example\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{time.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGrandom_ext}{mass extinction}, \protect\hyperlink{EGneg_ext}{negative trait extinction}, \protect\hyperlink{EGred_spec}{reducing speciation rate}, \protect\hyperlink{EG_change_trait}{changing trait process}, \protect\hyperlink{EG_change_modif}{changing modifiers}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{random.extinction}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGrandom_ext}{mass extinction}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{trait.extinction}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGneg_ext}{negative trait extinction}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{taxa.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGbg_ext}{background extinction}, \protect\hyperlink{EG_modify_brlen}{change branch length}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{bd.params.update}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGbg_ext}{background extinction}, \protect\hyperlink{EGred_spec}{reducing speciation rate}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{traits.update}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_trait}{changing trait process}, \protect\hyperlink{EG_change_correlation}{changing trait correlation}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{trait.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_correlation}{changing trait correlation}, \href{EG_founding_traits}{generating a subtree with a different process}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{modifiers.update}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_modif}{changing modifiers}, \protect\hyperlink{EG_modify_brlen}{change branch length}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{parent.traits}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_modif}{changing modifiers}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{founding.event}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_founding_purebirth}{generating a subtree with no extinction}, \href{EG_founding_traits}{generating a subtree with a different process}\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

We will then create a event that triggers when reaching half of the simulation time (using \texttt{time.condition(2.5)}).
This event will target \texttt{"taxa"}, i.e.~the number of species, and randomly remove 80\% of them (using \texttt{random.extinction(0.8)}):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# 80\% mass extinction at time 4}
\NormalTok{mass.extinction \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                    \DataTypeTok{condition =} \KeywordTok{time.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                    \DataTypeTok{target =} \StringTok{"taxa"}\NormalTok{,}
                    \DataTypeTok{modification =} \KeywordTok{random.extinction}\NormalTok{(}\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

And our extinction event which triggers after reaching time 4 (\texttt{time.condition(4)}), targets the \texttt{"taxa"} and modifies the extinction for species with traits lower than 0.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Extinction of any tips with trait \textless{} 1 at time 4}
\NormalTok{trait.extinction \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                      \DataTypeTok{target =} \StringTok{"taxa"}\NormalTok{,}
                      \DataTypeTok{condition =} \KeywordTok{time.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                      \DataTypeTok{modification =} \KeywordTok{trait.extinction}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{,}
                                                      \DataTypeTok{condition =} \StringTok{\textasciigrave{}}\DataTypeTok{\textless{}}\StringTok{\textasciigrave{}}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

And our extinction event which triggers after reaching time 4 (\texttt{time.condition(4)}), targets the \texttt{"bd.params"} and modifies the speciation rate to 1/3:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Reducing speciation after reaching time 4}
\NormalTok{reduced.speciation \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                      \DataTypeTok{condition =} \KeywordTok{time.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                      \DataTypeTok{target =} \StringTok{"bd.params"}\NormalTok{,}
                      \DataTypeTok{modification =} \KeywordTok{bd.params.update}\NormalTok{(}\DataTypeTok{speciation =} \DecValTok{1}\OperatorTok{/}\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

And our event which triggers after reaching time 5 (\texttt{time.condition(5)}), targets the \texttt{"traits"} and modifies the process to \texttt{OU.process}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Create an event to change the trait process}
\NormalTok{change.process.to.OU \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                  \DataTypeTok{condition    =} \KeywordTok{time.condition}\NormalTok{(}\DecValTok{5}\NormalTok{),}
                  \DataTypeTok{target       =} \StringTok{"traits"}\NormalTok{,}
                  \DataTypeTok{modification =} \KeywordTok{traits.update}\NormalTok{(}\DataTypeTok{process =}\NormalTok{ OU.process))}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmptzertD/file14f8c465f7d38}

@@

And our extinction event which triggers after reaching time 3 (\texttt{time.condition(3)}), targets the \texttt{"modifiers"} and modifies birth-death rule as follows:
* When a species is descendant from a parent with a negative trait value (\texttt{negative.trait.condition});
* Then increase your chances of going extinct by +1 (\texttt{increase.extinction.1})

@@

\CommentTok{\#\# Update the modifier}
\NormalTok{change.speciation \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
    \DataTypeTok{condition    =} \KeywordTok{time.condition}\NormalTok{(}\DecValTok{3}\NormalTok{),}
    \DataTypeTok{target       =} \StringTok{"modifiers"}\NormalTok{,}
    \DataTypeTok{modification =} \KeywordTok{modifiers.update}\NormalTok{(}\DataTypeTok{speciation =}\NormalTok{ speciation,}
                                    \DataTypeTok{condition  =}\NormalTok{ negative.trait.condition,}
                                    \DataTypeTok{modify     =}\NormalTok{ increase.extinction}\FloatTok{.1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

>>>>

\title{treats manual}
\author{Thomas Guillerme (\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}})}
\date{2023-11-22}

@@

@@

\begin{verbatim}
## $main
## $main$A
## $main$A$process
## function (x0 = 0, edge.length = 1, Sigma = diag(length(x0)), 
##     ...) 
## {
##     return(t(MASS::mvrnorm(n = 1, mu = x0, Sigma = Sigma * edge.length, 
##         ...)))
## }
## <bytecode: 0x559776bed260>
## <environment: namespace:treats>
## 
## $main$A$start
## [1] 0
## 
## $main$A$trait_id
## [1] 1
## 
## 
## 
## $background
## NULL
\end{verbatim}

@@

\begin{verbatim}
## $waiting
## $waiting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(rexp(1, sum(lineage$n * (bd.params$speciation + bd.params$extinction))))
## }
## <bytecode: 0x559775650390>
## <environment: namespace:treats>
## 
## $waiting$internal
## NULL
## 
## 
## $selecting
## $selecting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(sample(lineage$n, 1))
## }
## <bytecode: 0x559775658740>
## <environment: namespace:treats>
## 
## $selecting$internal
## NULL
## 
## 
## $speciating
## $speciating$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(runif(1) <= (bd.params$speciation/(bd.params$speciation + 
##         bd.params$extinction)))
## }
## <bytecode: 0x559775657e10>
## <environment: namespace:treats>
## 
## $speciating$internal
## NULL
## 
## 
## $call
## $call$waiting
## $call$waiting$fun
## [1] "default"
## 
## 
## $call$selecting
## $call$selecting$fun
## [1] "default"
## 
## 
## $call$speciating
## $call$speciating$fun
## [1] "default"
\end{verbatim}

@@

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.19\columnwidth}\raggedright
function\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright
arguments\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright
input\strut
\end{minipage} & \begin{minipage}[b]{0.33\columnwidth}\raggedright
what does it do\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.19\columnwidth}\raggedright
\texttt{age.condition}, \texttt{taxa.condition}, \texttt{trait.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
x\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
the value to check the target against\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
e.g.~time = x, taxa = x or trait = x (see below)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright
condition\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
a relational operator\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
e.g.~\texttt{==}, \texttt{\textgreater{}}, \texttt{!=}, etc. for asking for example \texttt{time\ ==\ x}\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.19\columnwidth}\raggedright
\ldots{}\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright
any additional arguments to be passed to the event\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright
\strut
\end{minipage} & \begin{minipage}[t]{0.33\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{itemize}
\tightlist
\item
  \texttt{age.condition} will trigger the \texttt{event} once a certain time is reached in the simulations and is the simplest/most basic condition with no arguments other than the time required. For example \texttt{age.condition(4,\ condition\ =\ \textasciigrave{}==\textasciigrave{})} will trigger the \texttt{event} once the simulations reaches 4 time units. Easy.
\item
  \texttt{taxa.condition} will trigger the \texttt{event} once a certain number of taxa are reached. This can be considered including or excluding fossil species. For example \texttt{taxa.condition(42,\ condition\ =\ \textasciigrave{}\textgreater{}=\textasciigrave{},\ living\ =\ TRUE)} will trigger the \texttt{event} if there are at least 42 living taxa.
\item
  \texttt{trait.condition} will trigger the \texttt{event} once a certain trait value is reached. This function allows to say which trait(s) to target (by default, the first one using \texttt{trait\ =\ 1}), what value of the trait to target (by default \texttt{what\ =\ max}) and whether to use and absolute trait value or not (\texttt{absolute\ =\ TRUE}). For example \texttt{trait.condition(1/3,\ condition\ =\ \textasciigrave{}\textgreater{}\textasciigrave{},\ trait\ =\ 1,\ what\ =\ sd)} will trigger the condition after the standard deviation of the first trait reaches 1/3.
\end{itemize}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# 80\% mass extinction at time 4}
\NormalTok{mass\_extinction \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                      \DataTypeTok{target =} \StringTok{"taxa"}\NormalTok{,}
                      \DataTypeTok{condition =} \KeywordTok{age.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                      \DataTypeTok{modification =} \KeywordTok{random.extinction}\NormalTok{(}\FloatTok{0.8}\NormalTok{))}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpqXtl0Y/file8c965577e651}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpqXtl0Y/file8c9618f635d6}

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.55\columnwidth}\raggedright
Functionality showcase\strut
\end{minipage} & \begin{minipage}[b]{0.39\columnwidth}\raggedright
Example\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{age.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGrandom_ext}{mass extinction}, \protect\hyperlink{EGneg_ext}{negative trait extinction}, \protect\hyperlink{EGred_spec}{reducing speciation rate}, \protect\hyperlink{EG_change_trait}{changing trait process}, \protect\hyperlink{EG_change_modif}{changing modifiers}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{random.extinction}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGrandom_ext}{mass extinction}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{trait.extinction}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGneg_ext}{negative trait extinction}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{taxa.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGbg_ext}{background extinction}, \protect\hyperlink{EG_modify_brlen}{change branch length}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{bd.params.update}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EGbg_ext}{background extinction}, \protect\hyperlink{EGred_spec}{reducing speciation rate}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{traits.update}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_trait}{changing trait process}, \protect\hyperlink{EG_change_correlation}{changing trait correlation}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{trait.condition}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_correlation}{changing trait correlation}, \href{EG_founding_traits}{generating a subtree with a different process}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{modifiers.update}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_modif}{changing modifiers}, \protect\hyperlink{EG_modify_brlen}{change branch length}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{parent.traits}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_change_modif}{changing modifiers}\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.55\columnwidth}\raggedright
\texttt{founding.event}\strut
\end{minipage} & \begin{minipage}[t]{0.39\columnwidth}\raggedright
\protect\hyperlink{EG_founding_purebirth}{generating a subtree with no extinction}, \href{EG_founding_traits}{generating a subtree with a different process}\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

We will then create a event that triggers when reaching half of the simulation time (using \texttt{age.condition(2.5)}).
This event will target \texttt{"taxa"}, i.e.~the number of species, and randomly remove 80\% of them (using \texttt{random.extinction(0.8)}):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# 80\% mass extinction at time 4}
\NormalTok{mass.extinction \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                    \DataTypeTok{condition =} \KeywordTok{age.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                    \DataTypeTok{target =} \StringTok{"taxa"}\NormalTok{,}
                    \DataTypeTok{modification =} \KeywordTok{random.extinction}\NormalTok{(}\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

And our extinction event which triggers after reaching time 4 (\texttt{age.condition(4)}), targets the \texttt{"taxa"} and modifies the extinction for species with traits lower than 0.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Extinction of any tips with trait \textless{} 1 at time 4}
\NormalTok{trait.extinction \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                      \DataTypeTok{target =} \StringTok{"taxa"}\NormalTok{,}
                      \DataTypeTok{condition =} \KeywordTok{age.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                      \DataTypeTok{modification =} \KeywordTok{trait.extinction}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{,}
                                                      \DataTypeTok{condition =} \StringTok{\textasciigrave{}}\DataTypeTok{\textless{}}\StringTok{\textasciigrave{}}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

And our extinction event which triggers after reaching time 4 (\texttt{age.condition(4)}), targets the \texttt{"bd.params"} and modifies the speciation rate to 1/3:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Reducing speciation after reaching time 4}
\NormalTok{reduced.speciation \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                      \DataTypeTok{condition =} \KeywordTok{age.condition}\NormalTok{(}\DecValTok{4}\NormalTok{),}
                      \DataTypeTok{target =} \StringTok{"bd.params"}\NormalTok{,}
                      \DataTypeTok{modification =} \KeywordTok{bd.params.update}\NormalTok{(}\DataTypeTok{speciation =} \DecValTok{1}\OperatorTok{/}\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

And our event which triggers after reaching time 5 (\texttt{age.condition(5)}), targets the \texttt{"traits"} and modifies the process to \texttt{OU.process}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Create an event to change the trait process}
\NormalTok{change.process.to.OU \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
                  \DataTypeTok{condition    =} \KeywordTok{age.condition}\NormalTok{(}\DecValTok{5}\NormalTok{),}
                  \DataTypeTok{target       =} \StringTok{"traits"}\NormalTok{,}
                  \DataTypeTok{modification =} \KeywordTok{traits.update}\NormalTok{(}\DataTypeTok{process =}\NormalTok{ OU.process))}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpqXtl0Y/file8c9623bb5976}

@@

And our extinction event which triggers after reaching time 3 (\texttt{age.condition(3)}), targets the \texttt{"modifiers"} and modifies birth-death rule as follows:
* When a species is descendant from a parent with a negative trait value (\texttt{negative.trait.condition});
* Then increase your chances of going extinct by +1 (\texttt{increase.extinction.1})

@@

\CommentTok{\#\# Update the modifier}
\NormalTok{change.speciation \textless{}{-}}\StringTok{ }\KeywordTok{make.events}\NormalTok{(}
    \DataTypeTok{condition    =} \KeywordTok{age.condition}\NormalTok{(}\DecValTok{3}\NormalTok{),}
    \DataTypeTok{target       =} \StringTok{"modifiers"}\NormalTok{,}
    \DataTypeTok{modification =} \KeywordTok{modifiers.update}\NormalTok{(}\DataTypeTok{speciation =}\NormalTok{ speciation,}
                                    \DataTypeTok{condition  =}\NormalTok{ negative.trait.condition,}
                                    \DataTypeTok{modify     =}\NormalTok{ increase.extinction}\FloatTok{.1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

### 2023-11-22 inst/treats_paper.Rmd
change time.condition to age.condition
For bypassing s3 issues on the CRAN

```{r, eval = FALSE}
## Creating an extinction that removes species with positive trait values
positive_extinction <- make.events(
    target       = "taxa",
    condition    = time.condition(15),
    modification = trait.extinction(x = 0, condition = `>=`))
```

For this event, the target is the number of taxa in the simulations.
This is indicated using the `target = "taxa"` argument.
Then the event is triggered using the argument `time.condition(15)` and modifies the internal `lineage` object using the ``trait.extinction(x = 0, condition = `>=`)`` function.
`time.condition` and `trait.extinction` are both function factories which are not gonna be detailed here (see @wickham2019advanced).
Effectively these arguments can be passed directly as standard functions.
For example, to trigger the event when reaching time 15 we can use the following function:

@@

```{r, message = FALSE}
## Creating a random mass extinction
random_extinction <- make.events(
    target       = "taxa",
    condition    = time.condition(15),
    modification = random.extinction(0.75))
## Creating an extinction that removes species with positive trait values
positive_extinction <- make.events(
    target = "taxa",
    condition = time.condition(15),
    modification = trait.extinction(x = 0, condition = `>=`))

>>>>

```{r, eval = FALSE}
## Creating an extinction that removes species with positive trait values
positive_extinction <- make.events(
    target       = "taxa",
    condition    = age.condition(15),
    modification = trait.extinction(x = 0, condition = `>=`))
```

For this event, the target is the number of taxa in the simulations.
This is indicated using the `target = "taxa"` argument.
Then the event is triggered using the argument `age.condition(15)` and modifies the internal `lineage` object using the ``trait.extinction(x = 0, condition = `>=`)`` function.
`age.condition` and `trait.extinction` are both function factories which are not gonna be detailed here (see @wickham2019advanced).
Effectively these arguments can be passed directly as standard functions.
For example, to trigger the event when reaching time 15 we can use the following function:

@@

```{r, message = FALSE}
## Creating a random mass extinction
random_extinction <- make.events(
    target       = "taxa",
    condition    = age.condition(15),
    modification = random.extinction(0.75))
## Creating an extinction that removes species with positive trait values
positive_extinction <- make.events(
    target = "taxa",
    condition = age.condition(15),
    modification = trait.extinction(x = 0, condition = `>=`))
